### 2021-09-06 input/experiments.tex
updated three-phase purge, small changes in implem, two column algo, smaller expe figs: 17 pages

\begin{figure}
  \centering\includegraphics[width=0.8\columnwidth]{img/as_cast_complexity.eps}
  \caption{\label{fig:complexity}Overhead of consistent partitioning
    using \NAME.}
  %% \caption{\label{fig:complexity}Cumulative average number of messages
  %%   per process in a network comprising 10k processes; along with the
  %%   duration before reaching consistent partitioning after adding or
  %%   removing a partition, totaling 100 new partition then removed.}
\end{figure}

@@

\begin{figure}
  \centering\includegraphics[width=0.8\columnwidth]{img/as_cast_geant.eps}
  \caption{\label{fig:geant}Overhead of \NAME in the partitioning of 2 clusters.}
\end{figure}

>>>>

%% \begin{figure*}
%%   \begin{center}
%%     \subfloat[Part A][\label{fig:complexity}Overhead of consistent partitioning
%%       using \NAME]
%%              {\includegraphics[width=0.5\columnwidth]{img/as_cast_complexity.eps}}
%%              \hfill
%%     \subfloat[Part B][\label{fig:geant}Overhead of \NAME in the partitioning of 2 clusters.]
%%              {\includegraphics[width=0.5\columnwidth]{img/as_cast_geant.eps}}
%%   \end{center}
%% \end{figure*}

\begin{figure}
  \centering\includegraphics[width=0.65\columnwidth]{img/as_cast_complexity.eps}
  \caption{\label{fig:complexity}Overhead of consistent partitioning
    using \NAME.}

  %% \caption{\label{fig:complexity}Cumulative average number of messages
  %%   per process in a network comprising 10k processes; along with the
  %%   duration before reaching consistent partitioning after adding or
  %%   removing a partition, totaling 100 new partition then removed.}
\end{figure}

@@

\begin{figure}
  \centering\includegraphics[width=0.65\columnwidth]{img/as_cast_geant.eps}
  \caption{\label{fig:geant}Overhead of \NAME in the partitioning of 2 clusters.}
\end{figure}

### 2021-09-06 input/figproofB.tex
updated three-phase purge, small changes in implem, two column algo, smaller expe figs: 17 pages

\begin{tikzpicture}[scale=0.9]

@@

  \newcommand\X{115pt}
  \newcommand\Y{-60pt}

@@

  \draw[<->] (5pt , 0.4*\Y) --
  node[below, font=\small, align=center]{\textbf{II: detection}}
  (-5pt + \X, 0.4*\Y); %% IV

@@

  \draw[fill=white] (-1.5*\X, 0) node{$\bm{X}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (-\X, 0) node{$\bm{a}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);  
  \draw[fill=white] (0, 0) node[color=\PY]{$\bm{b}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (\X, 0) node[color=\PX]{$\bm{c}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (2*\X, 0) node[color=\PX]{$\bm{d}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (0.5*\X, 1.25*\Y)node{$\bm{Y}$}+(-5pt, -5pt) rectangle +(5pt, 5pt);

>>>>

\begin{tikzpicture}[scale=0.75]

@@

  \newcommand\X{120pt}
  \newcommand\Y{-60pt}

@@

  \draw[<->] (5pt , 0.4*\Y) --
  node[below, font=\small, align=center]
  {\textbf{II: detection}\\\textbf{of inconsistencies}}
  (-5pt + \X, 0.4*\Y); %% IV

@@

  \draw[fill=white] (-1.5*\X, 0) node{$\bm{X}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (-\X, 0) node{$\bm{a}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);  
  \draw[fill=white] (0, 0) node[color=\PY]{$\bm{b}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (\X, 0) node[color=\PX]{$\bm{c}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (2*\X, 0) node[color=\PX]{$\bm{d}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (0.5*\X, 1.25*\Y) node{$\bm{Y}$}+(-5pt, -5pt) rectangle +(5pt, 5pt);

### 2021-09-06 input/proposal.tex
updated three-phase purge, small changes in implem, two column algo, smaller expe figs: 17 pages

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[a, b, c, d]$ delivered
  and forwarded a message $\alpha_a$. Some \processes may have
  delivered messages about other partitions before or after
  $\alpha_a$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_a$:] \Processes such as
    \Process~$a$ where the last and best partition is still $P_a$. By
    propagating from \process to \process the corresponding $\delta_a$
    message, these \processes purge $\alpha_a$.
  \item [\processes with $\alpha_a^y \rightarrow \alpha_e^z$ with last
    $\alpha_e^z$, for $\alpha_e^z < \alpha_a^y$:] \Processes such as
    \Process~$e$ that do not directly suffer from the non-delivery of
    $\delta_a$. They need to purge their own $P_e$ if need
    be. Nonetheless, they deliver and forward $\alpha_e$ that at least
    one following \process will receive.
  \item [\processes with $\delta_e \rightarrow \alpha_d$:] \Processes
    such as \Process~$f$ that do not belong to the preceding category,
    for they already delivered $\delta_e$. Nevertheless, best eventual
    forwarding guarantees that they receive $\alpha_e$, which
    contradicts their history or state. In such a case, the detecting
    \processes purge their own $\alpha_a$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted. Indeed their parent may have
    blocked the corresponding $\delta$ messages and subsequent
    \processes must purge their possibly stale state, for the sake of
    consistency.  We note such a notification $\Delta_a$ to emphasize
    the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_a$ receiving $\Delta_a$ from
    their parent:] \Processes such as \Process~$g$ that have
    $\alpha_a$ but preceding delivered messages are not
    important. Such \processes must trust detecting \processes by
    delivering and forwarding $\Delta_a$. It is worth noting that
    these messages are subject to all aforementioned blocking
    conditions, but are also solved by aforementioned mechanisms.
  \end{asparadesc}

  Propagation~I' terminates: a \process does not deliver a $\delta$
  message if it has already delivered it. Propagation~IV terminates: a
  \process does not deliver a looping message.
  % ; and since \processes forward the messages they deliver, upstream
  % \processes eventually receive and may deliver messages detected as
  % issue. In Figure~\ref{fig:proofB}, \Process~$e$ eventually
  % receives $\delta_f$ from \Process~$b'$.
  These three phases propagation-detection-propagation ensure that
  \processes eventually purge stale $\alpha$ messages from the system,
  leaving room for Propagation~I.
  %
  %  \TODO{cannot loop add from $e$ , undo from $b'$ because $b'$ sent
  % delete to $e$, it will solve the inconsistency.}
  %
  %% In the meantime, \processes that still belong to a partition can
  %% propagate their respective last $\alpha$ message, and reach
  %% consistent partitioning.
\end{proof}

@@

\begin{algorithm}[h]
  \input{input/algoascast.tex}
  \caption{\label{algo:ascast}\NAME at \Process~$p$ in static networks.}
\end{algorithm}

To implement the removal of stale messages as stated in
Definition~\ref{def:lwo}, each \process maintains a vector of versions
that associates the respective known local counter of each known
source, or has-been source. Each message $m$ carries its source $s$
along with its version $c$ that we subscript $m_{s, c}$.  Each
\process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  Its size
increases linearly with the number of sources that the \process has
ever known, which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case.
%% Unfortunately, \processes cannot
%% remove any entry from this vector without running a costly distributed
%% consensus protocol to ensure that no \process can mistake an old but
%% removed entry for a new one.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such a
$\delta_{s, c}^{\pi}$ message implements the uncertain $\Delta$
message of Figure~\ref{fig:proofB}. It fulfils the double role of
removing stale control information without modifying the local counter
of the source (since the partition may still exist), and removing
loops to ensure termination (since \processes do not modify their
local state to reflect such a delivery, messages piggyback
monotonically such increasing control information).

$\delta_{s, c}$ messages used in Propagation~I' are more efficient
than $\delta_{s, c}^\pi$ messages used in Propagation~IV, both in
propagation time and traffic. First, the former propagate using every
communication link available in the partition, while the latter
propagate following the dissemination tree of the corresponding
$\alpha$ to purge. Second, the former only piggyback source and
counter while the latter piggyback the path taken by the corresponding
$\alpha$ to purge.  In the worst case, a path includes every \process
in the system but converges to the average diameter of
partitions. Nevertheless, message paths and \NAME synergyze with each
other: Paths tend to be smaller as the number of sources in the system
increases. Bloom filters could introduce another trade-off by further
decreasing the size of path data structures at the cost of premature
stopped propagation that could impair consistent
partitioning~\cite{whitaker2002forwarding}.

@@

\begin{algorithm}
  \input{input/algoedges.tex}
  \caption{\label{algo:edges}\NAME at \Process~$p$ in dynamic networks.}
\end{algorithm}

>>>>

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[a, b, c, d]$ delivered
  and forwarded a message $\alpha_X$. Some \processes may have
  delivered messages about other partitions before or after
  $\alpha_X$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_X$:] \Processes such as
    \Process~$a$ where the last and best partition is still $P_X$. By
    propagating from \process to \process the corresponding $\delta_X$
    message, these \processes purge $\alpha_X$.
  \item [\processes with $\alpha_X^x \rightarrow \alpha_Y^y$ with last
    $\alpha_Y^y$, for $\alpha_Y^y< \alpha_X^x$:] \Processes such as
    \Process~$b$ that do not directly suffer from the non-delivery of
    $\delta_X$. They need to purge their own $P_Y$ if need
    be. Nonetheless, they deliver and forward $\alpha_Y$ that at least
    one following \process will receive.
  \item [\processes with $\delta_Y \rightarrow \alpha_X$:] \Processes
    such as \Process~$c$ that do not belong to the preceding category,
    for they already delivered $\delta_Y$. Nevertheless, best eventual
    forwarding guarantees that they receive $\alpha_Y$, which
    contradicts their history or state. In such a case, the detecting
    \processes purge their own $\alpha_X$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted. Indeed their parent may have
    blocked the corresponding $\delta$ messages and subsequent
    \processes must purge their possibly stale state, for the sake of
    consistency.  We note such a notification $\Delta_X$ to emphasize
    their relative similarity to $\delta_X$ messages.
    % the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_X$ receiving $\Delta_X$ from
    their parent:] \Processes such as \Process~$d$ that have
    $\alpha_X$ but preceding delivered messages are not
    important. Such \processes must trust detecting \processes by
    delivering and forwarding $\Delta_X$. It is worth noting that
    these messages are subject to all aforementioned blocking
    conditions, but are also solved by aforementioned mechanisms.
  \end{asparadesc}

  Propagation~I terminates: a \process does not deliver a $\delta$
  message if it has already delivered it. Propagation~III terminates: a
  \process does not deliver a looping message.
  % ; and since \processes forward the messages they deliver, upstream
  % \processes eventually receive and may deliver messages detected as
  % issue. In Figure~\ref{fig:proofB}, \Process~$e$ eventually
  % receives $\delta_f$ from \Process~$b'$.
  These three phases propagation-detection-propagation ensure that
  \processes eventually purge stale $\alpha$ messages from the system,
  leaving room for Propagation~I.
  %
  %  \TODO{cannot loop add from $e$ , undo from $b'$ because $b'$ sent
  % delete to $e$, it will solve the inconsistency.}
  %
  %% In the meantime, \processes that still belong to a partition can
  %% propagate their respective last $\alpha$ message, and reach
  %% consistent partitioning.
\end{proof}

@@

\begin{algorithm}[h!]
  \input{input/algoascast.tex}
  \caption{\label{algo:ascast}\NAME at \Process~$p$ in static networks.}
\end{algorithm}

To implement the removal of stale messages as stated in
Definition~\ref{def:lwo}, each \process maintains a vector of versions
that associates the respective known local counter of each known
source, or has-been source. Each message $m$ carries its source $s$
along with its version $c$ that we subscript $m_{s, c}$.  Each
\process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
% This vector enables the simplest Propagation~I and Propagation~I'.
Its size increases linearly with the number of sources that the
\process has ever known, which is the number of \processes in the
system $\mathcal{O}(V)$ in the worst case.
%% Unfortunately, \processes cannot
%% remove any entry from this vector without running a costly distributed
%% consensus protocol to ensure that no \process can mistake an old but
%% removed entry for a new one.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\Delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such a
$\Delta_{s, c}^{\pi}$ fulfils the double role of removing stale
control information without modifying the local counter of the source
(since the partition may still exist), and removing loops to ensure
termination (since \processes do not modify their local state to
reflect such a delivery, messages piggyback monotonically such
increasing control information).

$\delta_{s, c}$ messages used in Propagation~I are more efficient than
$\Delta_{s, c}^\pi$ messages used in Propagation~III, both in
propagation time and traffic. First, the former propagate using every
communication link available in the partition, while the latter
propagate following the dissemination tree of the corresponding
$\alpha$ to purge. Second, the former only piggyback source and
counter while the latter piggyback the path taken by the corresponding
$\alpha$ to purge.  In the worst case, a path includes every \process
in the system but converges to the average diameter of
partitions. Nevertheless, message paths and \NAME synergyze with each
other: Paths tend to be smaller as the number of sources in the system
increases.
%% Bloom filters could introduce another trade-off by further
%% decreasing the size of path data structures at the cost of premature
%% stopped propagation that could impair consistent
%% partitioning~\cite{whitaker2002forwarding}.

@@

\begin{algorithm}[h!]
  \input{input/algoedges.tex}
  \caption{\label{algo:edges}\NAME at \Process~$p$ in dynamic networks.}
\end{algorithm}
